Inducing Domain Specific Languages for Bayesian Program Learning

### 2018-02-02 paper.tex
new figure explaining how the different components bootstrap off of each other

\usepackage{verbatim}
 \usepackage{booktabs}
 % For algorithms
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{tikz}
\usetikzlibrary{fit,bayesnet}
%\usetikzlibrary{arrows.meta}
\usetikzlibrary{positioning}
%\usetikzlibrary{decorations.text,decorations.pathreplacing}
%\usetikzlibrary{decorations.pathmorphing}
\usepackage{dsfont}
\usepackage{amsmath}
\usepackage{hyperref}
\DeclareMathOperator*{\argmin}{arg\,min} % thin space, limits underneath in displays
\DeclareMathOperator*{\argmax}{arg\,max} % thin space, limits underneath in displays
\DeclareMathOperator{\argmin}{argmin} % no space, limits underneath in displays

@@

  \node[latent] at (1,3) (d){$\mathcal{D}$};
  \node[latent] at (2.5,3) (t){$\theta$};
  \node[latent] at (1,1.5) (z){$p$};
  \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
  \node[obs] at (1,-0.5) (x) {$x$};
  \edge {z}{x};
  \edge {d,t}{z};
  \plate {}{(tx)(z)(x)}{$x\in X$};
  \draw [->,dashed] (x.east) to[out = 0,in = -90] node(nn)[fill = white]{
          \includegraphics[width = 1.5cm]{figures/MLP.png} 
  } (tx.south);
  \node at (nn.south) {$q(x)$};
  \draw [->,dashed] (d.west) to[out = 180,in = 180] (z.west);
  \draw [->,dashed] (tx.west) -- (z.east);
  \end{tikzpicture}
  \caption{DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives with probabilities $\theta$ (Algorithm~\ref{programGenerativeModel}). We observe program outputs $x$. A neural network $q(\cdot )$ called the \emph{recognition model} regresses from program outputs to a distribution over programs ($\theta^{(x)} = q(x)$). Solid arrows correspond to the top-down generative model. Dashed arrows correspond to the bottom-up recognition model.}\label{graphicalModel}
\end{figure}

@@

\noindent\textbf{Putting it all together.}
Algorithm~\ref{mainAlgorithm} describes how we combine the
program search, recognition model training, and DSL induction.

\begin{algorithm}[tb]
   \caption{The \system Algorithm}
   \label{mainAlgorithm}
   \begin{algorithmic}
     \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
     \STATE \textbf{Hyperparameters:} Maximum frontier size $F$
     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, bottom-up recognition model $q(\cdot)$
     \STATE Initialize $\mathcal{D}_0\gets \mathcal{D}$, $\theta_0\gets \text{uniform}$, $q_0(\cdot ) = \theta_0$
     \FOR{$i=1$ {\bfseries to} $I$}
     \FOR{$x:\tau\in X$}
     \STATE  $\mathcal{F}_x\gets \{z| z\in \text{enumerate}(\mathcal{D}_{i - 1},q_{i - 1}(x),F)\cup\text{enumerate}(\mathcal{D}_{i - 1},\theta_{i - 1},F) \text{ if }\probability[x|z] > 0\}$
     \ENDFOR
     \STATE $\mathcal{D}_i,\theta_i\gets $induceDSL$(\{\mathcal{F}_x\}_{x\in X})$
     %% \STATE Define $Q_x(z) \propto \begin{cases}
     %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
     %%   0&x\not \in \mathcal{F}_x
     %% \end{cases}

     \STATE $q_i\gets \text{train recognition model to maximize }\mathcal{L}_{\text{RM}}$
      \ENDFOR
 \STATE \textbf{return} $\mathcal{D}^I,\theta^I,q^I$
\end{algorithmic}
\end{algorithm}

>>>>

\usepackage{verbatim}
 \usepackage{booktabs}
 % For algorithms
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{tikz}
\usetikzlibrary{fit,bayesnet}
%\usetikzlibrary{arrows.meta}
%\usetikzlibrary{positioning}
%\usetikzlibrary{decorations.text}
%\usetikzlibrary{decorations.pathmorphing}
\usepackage{dsfont}
\usepackage{amsmath}
\usepackage{hyperref}
\DeclareMathOperator*{\argmin}{arg\,min} % thin space, limits underneath in displays
\DeclareMathOperator*{\argmax}{arg\,max} % thin space, limits underneath in displays
\DeclareMathOperator{\argmin}{argmin} % no space, limits underneath in displays

@@

  \node[latent] at (1,3) (d){$\mathcal{D}$};
  \node[latent] at (2.5,3) (t){$\theta$};
  \node[latent] at (1,1.5) (z){$p$};
  \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
  \node[obs] at (1,-0.5) (x) {$x$};
  \edge {z}{x};
  \edge {d,t}{z};
  \plate {}{(tx)(z)(x)}{$x\in X$};
  \draw [->,dashed] (x.east) to[out = 0,in = -90] node(nn)[fill = white]{
          \includegraphics[width = 1.5cm]{figures/MLP.png} 
  } (tx.south);
  \node at (nn.south) {$q(x)$};
  \draw [->,dashed] (d.west) to[out = 180,in = 180] (z.west);
  \draw [->,dashed] (tx.west) -- (z.east);

  \begin{scope}[shift = {(1,-1)}]
    \node[align = center](DSL) at (6,4) {DSL: $\mathcal{D}$};
    \node[align = center](synthesis) at (4,2) {Search for \\programs: $p$};
    \node[ align = center](recognitionModel) at (6,0) {Recognition \\model: $q$};

    \draw [->,thick] (DSL.west) to[out = 180,in = 90] node[above,rotate = 45]{{\footnotesize Makes tractable}} (synthesis.north);
    \draw [->,thick] (synthesis.east) to[out = 0,in = -90] node[above,rotate = 70]{{\footnotesize Trains}} (DSL.south);
    \draw [->,thick] (synthesis.east) to[out = 0,in = 90] node[below,rotate = -60]{{\footnotesize Trains}} (recognitionModel.north);
    \draw [->,thick] (recognitionModel.west) to[out = 180,in = -90] node[below,rotate = -70]{{\footnotesize Makes tractable}} (synthesis.south);
    \draw [->,thick] (DSL.east) to[out = 0,in = 0] node[above,rotate = -90]{{\footnotesize Trains}} (recognitionModel.east);
  \end{scope}
  \end{tikzpicture}
  \caption{DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives with probabilities $\theta$ (Algorithm~\ref{programGenerativeModel}). We observe program outputs $x$. A neural network $q(\cdot )$ called the \emph{recognition model} regresses from program outputs to a distribution over programs ($\theta^{(x)} = q(x)$). Solid arrows correspond to the top-down generative model. Dashed arrows correspond to the bottom-up recognition model.}\label{graphicalModel}
\end{figure}

@@

\subsection{Implementing the \system}

Algorithm~\ref{mainAlgorithm} describes how we combine the
program search, recognition model training, and DSL induction.
We additionally make the following implementation details:
(1) On the first iteration we do \emph{not} train the recognition model on samples from
the generative model because, on the first iteration, we have not yet learned a
new DSL, so we instead trained the network to only maximize $\mathcal{L}_{\text{AE}}$;
(2) Because the frontiers can grow very large,
we only keep around the top $10^4$ programs $p$ in each frontier $\mathcal{F}_x$
with the highest $\probability[x,p|\mathcal{D},\theta]$;
(3) During both DSL induction and neural net training, we calculate $\text{score}(\mathcal{D}')$ and $\mathcal{L}_{\text{RM}}$
by only summing over the top $K$
programs in $\mathcal{F}_x$.
We found that $K = 2$ sufficed.

\begin{algorithm}[tb]
   \caption{The \system Algorithm}
   \label{mainAlgorithm}
   \begin{algorithmic}
     \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
     \STATE \textbf{Hyperparameters:} Maximum frontier size $F$
     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
     \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
     \FOR{$i=1$ {\bfseries to} $I$}
%     \FOR{$x:\tau\in X$}
     \STATE  $\mathcal{F}^{(\theta)}_x\gets \{z| z\in \text{enumerate}(\mathcal{D},\theta,F)\text{ if }\probability[x|z] > 0\}$
 %    \ENDFOR
     \STATE $q_i\gets \text{train recognition model to maximize }\mathcal{L}_{\text{RM}}$
     \STATE  $\mathcal{F}^{(q)}_x\gets\{z|z\in \text{enumerate}(\mathcal{D},q(x),F)\text{ if }\probability[x|z] > 0\}$
     \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{(\theta)}_x\cup\mathcal{F}^{(q)}_x\}_{x\in X})$
     %% \STATE Define $Q_x(z) \propto \begin{cases}
     %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
     %%   0&x\not \in \mathcal{F}_x
     %% \end{cases}
      \ENDFOR
 \STATE \textbf{return} $\mathcal{D},\theta,q$
\end{algorithmic}
\end{algorithm}
